Sulfur Dioxide on Water's Surface

### 2011-06-23 DensityResults.tex
Rewrote some of draft 4, but still need to finish up results and conclusions section.
Draft 5 revisions - finished Geri's edits. Time for a re-read.

One measure of surface activity is the spatial distribution of molecules in the interfacial region. The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained in the bulk.

@@

In the simulated neat-water slab, the \suldiox~molecular density distribution indicates an affinity for the water surface. During the course of the simulation, the \suldiox~does not venture into the water bulk. Also, the \suldiox~does not escape the water surface into the gas phase, but remains located within 5-\angs~of the surface. This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS). The \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction. In a later section of this article we look at the effect of \suldiox~on surface water orientation.

The saturated surface simulation results indicate that under the high concentration conditions, \suldiox~accumulation at the surface is increased. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry. However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods, and is furthermore verified by our experimental results.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/density.png}
		\caption{Molecular density distributions of \wat~and \suldiox~calculated along the long-axis of the simulated cells. The position axis shows the distance of the given molecule from the top surface of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk. Distributions of both the neat-water simulation with a single \suldiox~(left, scaled 10x for clarity), and the saturated system (right) are shown.}
		\label{fig:density}
	\end{center}
\end{figure}

>>>>

One measure of surface activity is the spatial distribution of molecules in the interfacial region. The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations. The results presented in figure \ref{fig:density} show both the water and \suldiox~density distributions across the two simulated slabs. As shown, the \suldiox~exposed to the neat-water system remained at the water slab surface. The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained well into the bulk water.

@@

In the simulated neat-water slab, the \suldiox~molecular density distribution concentrates near the water surface, indicating an affinity for the interfacial region. This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS). In the experimental system, the \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction. In a later section of this article we look at the effect of \suldiox~on surface water orientation.  During the course of the simulation, the \suldiox~does not venture into the bulk water. Also, the \suldiox~does not escape the water surface into the gas phase, but remains located within 5-\angs~of the surface.

The saturated solution simulation results indicate that under the high concentration of \suldiox, the \suldiox~accumulation at the surface is increased. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry. However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/density.png}
		\caption{Molecular density distributions of \wat~(black) and \suldiox~(red) calculated along the long-axis of the simulated cells. The position axis shows the distance from the instantaneous surface location of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk. Distributions of both the neat-water simulation with a single \suldiox~(left, with the \suldiox~density scaled 10x for clarity), and the saturated system (right) are shown.}
		\label{fig:density}
	\end{center}
\end{figure}

### 2011-06-23 EquilibratedOrientation.tex
Draft 5 revisions - finished Geri's edits. Time for a re-read.

The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase. The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface. The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.

The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above the surface, and then the distribution becomes isotropic further into the water bulk of both systems. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface. Thus, both the neat-water and saturated systems have waters lying mostly flat to the plane of the interface at a distance of 0\angs.

>>>>

The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase. Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface. The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.

The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the surface, becoming progressively isotropic further into the water bulk of both systems. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface. Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface.

### 2011-06-23 Methods.tex
Draft 5 revisions - finished Geri's edits. Time for a re-read.

The neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration. The saturated system had 22 \suldiox~molecules in the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure. The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The second, more concentrated system examined consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M. This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface. Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.

@@

	In this portion of the study we create orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas. The reference point used for depth calculations is the location of the water surface. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.

@@

The new procedure described previously for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b. It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$. All depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding timestep of the trajectories.

>>>>

The low and high concentration simulated systems, ``neat-water'' and saturated, respectively, were created as follows: the ``neat-water'' simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true ``neat-water'' air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an \suldiox~gas pressure. The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The more concentrated ``saturated'' system consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M. This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface. Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates recorded every 100 fs.

@@

	In this portion of the study we create orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas. The reference point used for depth calculations is the location of the water surface. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.

	In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.

@@

Using this alternative method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b. It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$. Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding timestep of the trajectories.
